---
title: "Compadre Analysis"
author: "Paul Glaum as part of Simon et al 2022"
date: '2022-04-08'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Compadre plant matrix repository

COMPADRE is a global online repository for matrix population models (MPMs) and metadata on plants Please see <https://jonesor.github.io/Rcompadre/index.html> & <https://jonesor.github.io/CompadreGuides/user-guide.html> for in-depth discussion of COMPADRE install and usage.

## Install the stable release package from CRAN. Also load ggplot:

```{r packages, message=FALSE}

#install.packages("Rcompadre")
##Remote install from github with:
#remotes::install_github("jonesor/Rcompadre", build_vignettes = FALSE)
##
library(Rcompadre)
require(ggplot2)
require(gridExtra)
library(kableExtra)
library(leaflet)
```

## Get up to date version of COMPADRE

```{r fetchCompadre}
compadre <- cdb_fetch("compadre")
```

### Stage No. distribution
Check the current distribution of different matrix structures in COMPADRE. We only show models made up of 10 stages or less:
```{r matDim, echo=FALSE}
table(compadre$MatrixDimension)[1:10]
```

## Check the breakdown of 3-stage plant models.
### Species
Check number of taxonomic species currently represented by 3-stage plant models:
```{r checkSpecies, echo=FALSE}
##Creating full species name column###
compadre$fullSpName=paste(compadre$Genus,compadre$Species,sep=" ")
length(unique(compadre$fullSpName[compadre$MatrixDimension==3]))
#107
```
### Genera
Check distribution of taxonomic genera currently represented by 3-stage plant models:
<div align="center">
```{r genera, results="asis", echo=FALSE}
#print("Check number of taxonomic genera currently represented by 3-stage plant models:")
GeneraNum=length(unique(compadre$Genus[compadre$MatrixDimension==3]));
#print("Breakdown across genera:")
knitr::kable(table(compadre$Genus[compadre$MatrixDimension==3]),
             col.names = c("Plant Genus","Freq in Database"),
             caption = paste("Breakdown across the",GeneraNum, "plant genera") )  %>% 
  kable_styling("striped", full_width = F) %>% 
 scroll_box(width = "500px", height = "500px")
``` 
</div>

### Families
Check distribution of taxonomic families currently represented by 3-stage plant models:
<div align="center">
```{r families, results="asis", echo=FALSE}
#print("Check number of taxonomic families currently represented by 3-stage plant models:")
FamiliesNum=length(unique(compadre$Family[compadre$MatrixDimension==3]))
#print("Breakdown across families:")
knitr::kable(table(compadre$Family[compadre$MatrixDimension==3]),
             col.names = c("Plant Family","Freq in Database"),
             caption = paste("Breakdown across the",FamiliesNum, "plant families") ) %>% 
  kable_styling("striped", full_width = F) %>% 
 scroll_box(width = "500px", height = "500px")
```
</div>

### Geographic distribution:
Check COMPADRE website above to see Eco-region abbreviations.

```{r geo, message=FALSE, echo=FALSE}
t1=knitr::kable(
  table(compadre$Continent[compadre$MatrixDimension==3]),
             col.names = c("Continent","Freq in Database"),
             caption = paste("Breakdown across continents"));
t2=knitr::kable( 
  table(compadre$Ecoregion[compadre$MatrixDimension==3]),
             col.names = c("Eco-Region","Freq in Database"),
             caption = paste("Breakdown across Regions") );

t1 %>%
  kable_styling(full_width = FALSE, position = "float_left")
t2 %>%
  kable_styling(full_width = FALSE, position = "left")


```

#### Map
```{r maps, message=FALSE}
##Create a subset of only 3-stage matrices
mat3=subset(compadre, MatrixDimension==3);
LLdf=data.frame(mat3$Lat,mat3$Lon,mat3$fullSpName);
colnames(LLdf)=c("lat","long","Species");
LLdf<- LLdf[complete.cases(LLdf), ];
LLdf<-LLdf[!duplicated(LLdf), ];

suppressWarnings(
  LLdf %>%
  leaflet( width = 900) %>%
  addTiles() %>%
  addMarkers(clusterOptions = markerClusterOptions(), popup = ~paste0(
      "<b><h4>","Species", "</h4></b><br>",LLdf$Species)  )
)

```

## Prepare demographic rates

Pull specific demographic rate data from the 3-stage models. Note, if you are running these script, this must be run before making figures below.

```{r, rates, message=FALSE}
##Create a subset of only 3-stage matrices
mat3=subset(compadre, MatrixDimension==3);
##COMPADRE uses "accessor functions" to obtain specific parts of the matrices.
##matF accesses the fecundity part of the matrix, akin to our rF
Fmats=matF(mat3);
##matU accesses the survival-related transitions, akin to our g12 and g2F
Umats=matU(mat3);

#To store entry. Fs for reproduction alone, F.U for reproduction & transitions. 
Fs=rep(); F.U=matrix(0,length(Fmats),4)
for(i in 1:length(Fmats) ) {
	Fs[i]=sum(Fmats[i][[1]][1,3] )
	F.U[i,1:3]=Umats[i][[1]][lower.tri(Umats[i][[1]])] ##only take the transitions
	F.U[i,4]=Fs[i] #include reproduction
}
#make easy to use data frame for our curated data. 
#name columns in the vein of our parameters. Note, d31 named for the 
#addition into stage 1 from stage 3 is the same as our rF
UnF=as.data.frame(F.U)
colnames(UnF)=c('g12','g13','g23','d31')
UnF=subset(UnF,UnF$d31>0&UnF$d31<100)
```

## Make plots

Scatter plots showing correlations between parameters:

```{r fig1, fig.height = 3, fig.width = 10, fig.align = "center", message=FALSE}
g12vrF=ggplot(UnF) + theme_bw() +
 geom_point(aes(x=g12,y=log(d31)),alpha=.4,size=2) +
 xlab(bquote('maturation rate'~(g['12'])) )+
 ylab(bquote('log(reproduction rate)'~(r[F])) )

g23vrF=ggplot(UnF) + theme_bw() +
 geom_point(aes(x=g23,y=log(d31)),alpha=.4,size=2) +
 xlab(bquote('maturation rate'~(g['2F'])) )+
 ylab(bquote('log(reproduction rate)'~(r[F])) )

g23vg12=ggplot(UnF) + theme_bw() +
 geom_point(aes(x=g23,y=g12),alpha=.4,size=2) +
 xlab(bquote('maturation rate'~(g['2F'])) )+
 ylab(bquote('germination rate'~(g['12'])) )

suppressWarnings(grid.arrange(g12vrF, g23vrF, g23vg12, ncol=3) )
```

Histograms of the value distribution for each parameter:

```{r fig2, fig.height = 3, fig.width = 10, fig.align = "center"}
Fss=Fs[Fs<100&Fs>0]

h1<-suppressWarnings(hist(na.omit(Fss),breaks=80,col='light blue',xlim=c(0,10.6),plot=FALSE));
h1$density = suppressWarnings(h1$counts/sum(h1$counts));
h2<-suppressWarnings(hist(na.omit(UnF$g12[UnF$g12>0]),breaks=40,col='light blue',plot=FALSE));
h2$density = suppressWarnings(h2$counts/sum(h2$counts));
h3<-suppressWarnings(hist(na.omit(UnF$g23[UnF$g23>0]),breaks=40,col='light blue',plot=FALSE));
h3$density = suppressWarnings(h3$counts/sum(h3$counts))

par( mfrow= c(1,3) )
plot(h1,freq=FALSE,col='light blue',xlim=c(0,10.6),
xlab="Measured Rate",ylab="Density",main="Reproduction into 1st Stage")#,ylim=c(0,.5))

plot(h2,freq=FALSE,col='light blue',
xlab="Measured Rate",ylab="Density",main="1st to 2nd Stage Transition")#,ylim=c(0,.5))

plot(h3,freq=FALSE,col='light blue',
xlab="Measured Rate",ylab="Density",main="2nd to 3rd Stage Transition")#,ylim=c(0,.5))

```
